Think deep for reef resilience

### 2025-08-17 paper/paper_draft.qmd
Expand on role of 10-15m depth range

Current reef restoration and intervention-investigation efforts on the GBR, such as
deployment of coral seeding and fragment devices, target reefs with suitable restoration
sites in the 1-10m depth range [@howlett2021; @randall2021; @whitman2024]. While there are
no explicit justifications for this depth range, restoration depth is likely constrained by
snorkeling and diving logistical accessibility, visual identification of sites from the
surface, in combination with tourism return value. Our results suggest that of the assessed
reefs ocurring within the 1-10m depth range (1736 reefs), 62% (1082) of reefs fall into the
low and medium clusters (across all GCMs at the GBRMPA management region scale), with many
unable to maintain positive carbonate budgets under future conditions. Partial dependence
analysis revealed a range of depths from 10-15m where the probability of a reef being
clustered into the High cover category increased substatially. Considering this 10-15m depth
range when assessing possible restoration sites may enhance restoration success, while still
maintaining effective logistical access to sites. Reef intervention planning is a complex
and challenging decision making process [@anthony2020], and our results suggest that
deployment site depth may be an important factor to additionally consider when assessing the
long-term resilience of deployment sites.

>>>>

Current reef restoration and intervention-investigation efforts on the GBR, such as
deployment of coral seeding and fragment devices, target reefs with suitable restoration
sites in the 1-10m depth range [@howlett2021; @randall2021; @whitman2024]. While there are
no explicit justifications for this depth range, restoration depth is likely constrained by
snorkeling and diving logistical accessibility, visual identification of sites from the
surface, in combination with tourism return value. Our results suggest that of the assessed
reefs ocurring within the 1-10m depth range (1736 reefs), 62% (1082) of reefs fall into the
low and medium clusters (across all GCMs at the GBRMPA management region scale), with many
unable to maintain positive carbonate budgets under future conditions. Partial dependence
analysis revealed a range of depths from 10-15m where the probability of a reef being
clustered into the High cover category increased substatially. Considering this 10-15m depth
range when assessing possible restoration sites may reduce the exposure of planted and
seeeded corals to high thermal stress, while still maintaining effective logistical access
to sites. Reef intervention planning is a complex and challenging decision making process
[@anthony2020], and our results suggest that deployment site depth may be an important
factor to additionally consider when assessing the long-term resilience of deployment sites.
